---
title: "Titanic Tutorial - DataScienceDojo"
author: "Darrell Mayson"
output:
  html_document:
    df_print: paged
---
## Resource: 'ggplot2 - Elegent Graphics for Data Analysis (book), 2nd Ed.'
### by Hadley Wickham (pub: Springer)


```{r setup, echo=FALSE, message=FALSE}
library(titanic)
library(ggplot2)
library(dplyr)
```

```{r create-df}
titanic <- read.csv("titanic.csv", stringsAsFactors = FALSE)
# Checking the structure
str(titanic)
# View(titanic) # commented as causes crash when knitting to HTML
```



### Setup as factors to enable plot

```{r setup-factors}
titanic$Pclass <- as.factor(titanic$Pclass)
titanic$Survived <- as.factor(titanic$Survived)
titanic$Sex <- as.factor(titanic$Sex)
titanic$Embarked <- as.factor(titanic$Embarked)
```

### Initial analysis and cleanup data (e.g., missing values) - [TODO]
```{r cleanup}
colSums(is.na(titanic))

```




## First question - What was the survival rate?
### As 'Survived' is now a factor (i.e., Categorical) variable, a bar-chart is a great visualisation to use.
```{r bar-chart-survival}
ggplot(titanic, aes(x = Survived)) +
  geom_bar()
```
### Add some plot customisations and a theme
```{r bar-chart-survival-improved}
ggplot(titanic, aes(x = Survived)) +
  theme_classic() +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates")
# Just to see what the actual percentages are for Survival (in console)
prop.table(table(titanic$Survived))
```

## What was the survival rate by Gender
We can use color to look at two aspects (i.e., dimensions) of the data simultaneously.

```{r survival-rate-by-gender, echo=FALSE}
ggplot(titanic, aes(x = Sex, fill = Survived)) +
  theme_classic() +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates by Gender")
```

## What was the survival rate by ticket class?

```{r survival-rate-by-ticket-class}
ggplot(titanic, aes(x = Pclass, fill = Survived)) +
  theme_bw() +
  geom_bar() +
  labs(y = "Passenger Count", x = "Class of Ticket",
       title = "Titanic Survival Rates by Ticket Class")
```
## OK, so what was the survival rate by GENDER by PCLASS (i.e., sex and ticket class)?
### We can leverage facets to further segment the data and enable 'visual-drill-down' into the data.
```{r survival-rate-by-sex-facet-by-pclass}
ggplot(titanic, aes(x = Sex, fill = Survived)) +
  theme_bw() +
  facet_wrap(~ Pclass) +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates by Sex & Ticket Class")
```
## What was the distribution by passenger ages?
```{r dist-by-passenger-age}
ggplot(titanic, aes(x = age)) +
  theme_bw() +
  geom_histogram(bins = 5) +
  labs(y = "Passenger Count", x = "Age (binwidth = 5",
       title = "Titanic Age Distribution")
```

